Topological approach to finding blackholes in directed networks

### 2019-12-02 blackholes/pavt/article.tex
Minor modification
Minor modifications

\begin{abstract}
In this paper we consider the task of finding so called Blackhole pattern in directed unweighed graphs.
Firstly, we analyse already existing algorithms and approaches. Secondly, we describe how structure of a graph
affects their efficiency. Then we introduce our approach to graph preprocessing. Finally, we describe topological sort based heuristic
and provide results of experimental comparison between our approach and previously developed algorithm.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{Directed networks $\cdot$ Subgraph mining}
\end{abstract}
%
\section{Blackhole search in directed graph}
%

@@

\FIXME{Хорошо бы ссылку на graph condensation}

As it was described in the Known Issues section, large SCCs waste lots of computational time without adding any new blackholes. 
For certain graph types (RMAT, SmallWorld) it can be crucial, because one SCC can contain up to 100\% of all graph's vertices. 
In such a case, it would be nice to consider large SCC as a single vertex, aggregating all incoming and outcoming peripheral edges
of original SCC.
On the preprocessing stage of the algorithm we reduce every SCC to a single vertex. The algorithm is not new and known as a graph condensation.

@@

In this article we use the following algorithm for finding graph condensation.
It was independently proposed by Kosaraju and Sharir back in 1979.\par 
First, we find topological sort of the initial graph. It takes O(E + V).
\begin{definition}
TODO: Topological sort
\end{definition}

\begin{definition}
Closure of vertex v is a set of all nodes reachable from v, including v.
\end{definition}

@@

When we finally have an array, containing graph vertices in the topologically sorted order, we do the following:

Define global array of used vertices. Initially, all vertices are unused.
Then for every vertex from bottom to top:
\begin{enumerate}
    \item Acquire closure of unused vertices reachable from current vertex; 
    \item Mark every vertex in the closure as used;
    \item All the acquired vertices belong to the same SCC.
\end{enumerate}

The fact given in the last step can be shortly explained in the following way:
topological sort guarantees that any vertex has index larger than indeces of her children.
That is why, given that we process vertices from bottom to top, by the moment we process the vertex, all it's children are already marked used and no longer available for search.
As all the ways down are prohibited and there's no way up, we will acquire an SCC.\par

As a result of the algorithm, described above, we have each vertex associated with certain SCC.
SCCs itself form a graph with number of vertices less than or equal to one of the initial graph.

@@

%
\subsection{Blackholes search}
When the graph is prepared, we can move on to searching blackholes. The task is combinatorial in its nature, therefore, we aim to
decrease the number of potential black holes.
To understand the issue, we first describe BruteForce approach. \par

@@

To BruteForce the task we iterate over all possible unordered sets of nodes. 
Then for every node we acquire its closure. Union of all such closures is a candidate blackhole.
If candidate blackhole is a weakly connected subgraph, then it is a blackhole.

With BruteForce algorithm it is possible to find the same blackhole several times.
We aim to avoid at least some of the duplicates and introduce heuristic for it.

According to the definitions above if some node is not a root of blackhole, it can be omitted, as set of roots
uniquely identify a blackhole.
Every non-root node in blackhole is reachable from one of the roots. Therefore, if we know reachability matrix for
the given graph, we can skip bloated non-basis combinations of nodes.
Of course, we could straightforward build the reachability matrix. But, it takes O(V \^ 3) time. Even though we decreased graph size
in preprocessing step, it still can result in lots of computation. That is why, we will only try to partially skip duplicates and the rest will
be filtered later.\par

>>>>

\begin{abstract}
In this paper we consider the task of finding so called Blackhole pattern in directed unweighed graphs.
Firstly, we analyse already existing algorithms and approaches. Secondly, we describe how structure of a graph
affects their efficiency. Then we introduce our approach to graph preprocessing. Finally, we describe topological sort based heuristic
and provide results of experimental comparison between our approach and previously developed algorithm.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{Directed networks $\cdot$ Subgraph mining}
\end{abstract}
%
\section{Blackhole search in directed graph}
applications Anti-Money Laundering \cite{semenov2017survey}

@@

As it was described in the Known Issues section, large SCCs waste lots of computational time without adding any new blackholes. 
For certain graph types (RMAT, SmallWorld) it can be crucial, because one SCC can contain up to 100\% of all graph's nodes. 
In such a case, it would be nice to consider large SCC as a single node, aggregating all incoming and outcoming peripheral edges
of original SCC.
On the preprocessing stage of the algorithm we reduce every SCC to a single node. The algorithm is not new and known as a graph condensation.

@@

In this article we use the following algorithm for finding graph condensation.
It was independently proposed by Kosaraju and Sharir back in 1979. \FIXME{Ссылку на эту статью} \par

First, we find topological sort of the initial graph. It takes O(E + V).
\begin{definition}
TODO: Topological sort
\end{definition}

\begin{definition}
Closure of node v is a set of all nodes reachable from v, including v.
\end{definition}

@@

When we finally have an array, containing graph nodes in the topologically sorted order, we do the following:

Define global array of used nodes. Initially, all nodes are unused.
Then for every node from bottom to top:
\begin{enumerate}
    \item Acquire closure of unused nodes reachable from current node; 
    \item Mark every node in the closure as used;
    \item All the acquired nodes belong to the same SCC.
\end{enumerate}

The fact given in the last step can be shortly explained in the following way:
topological sort guarantees that any node has index larger than indeces of her children.
That is why, given that we process nodes from bottom to top, by the moment we process the node, all it's children are already marked used and no longer available for search.
As all the ways down are prohibited and there's no way up, we will acquire an SCC.\par

As a result of the algorithm, described above, we have each node associated with certain SCC.
SCCs itself form a graph with number of nodes less than or equal to one of the initial graph.

@@

%
\subsection{Blackholes search}
When the graph is prepared, we can move on to searching blackholes. The problem is combinatorial in its nature, therefore, we aim to
decrease the number of potential blackholes.
To understand the issue, we first describe a brute-force approach.

@@

%TODO: осознать
In the BruteForce algorithm we iterate over all possible unordered sets of nodes. 
Then for every node we acquire its closure. Union of all such closures is a potential blackhole.
If the candidate blackhole is a weakly connected subgraph, then it is a blackhole.

With the BruteForce algorithm it is possible to find the same blackhole several times.
We aim to avoid at least some of the duplicates and introduce heuristic for it.

According to the definitions above if some node is not a root of a blackhole, it can be omitted, as set of roots uniquely identify a blackhole.
Every non-root node in a blackhole is reachable from one of the roots. Therefore, if we know reachability matrix for a given graph, we can skip bloated non-basis combinations of nodes.
Of course, we could straightforward build the reachability matrix. But, it takes O(V \^ 3) time. Even though we decreased graph size
in preprocessing step, it still can result in lots of computation. That is why, we will only try to partially skip duplicates and the rest will
be filtered later.\par
